---
title: Iris Sepal Width ANOVA
author: D. Zeitler
date: Sys.Date()
output: pdf_document
---

*First get a few setup things out of the way*
```{r setup,echo=FALSE}
library(tidyverse)
library(mosaic)
#library(RcmdrMisc)
library(multcomp, pos=4)
library(abind, pos=4)
library(MASS, pos=4)
data(iris, package="datasets")
```
Note that the order here may be important. There are functions that get modified by multiple packages. The result seems ok, but might cause some problems.

# Parametric one-way testing

We're going to use the built-in Iris data set to show and example of an analysis of variance.

Start by looking at box plots. We're looking for similar box sizes and symmetry in each of the groups. We'd like normality if  possible. 

```{r}
tally(~Species,data=iris)
bwplot(Sepal.Width~Species, data=iris)
gf_qq(~Sepal.Width|Species, data=iris) %>%
gf_qqline(~Sepal.Width|Species, data=iris)
favstats(Sepal.Width~Species, data=iris)
```

```{r}
histogram(~Petal.Length|Species,data=iris)
qqmath(~Petal.Length|Species,data=iris)
```

We could do three t-tests for all pairs of species.

```{r}
iris %>%
  filter(Species!='virginica') %>%
  t.test(Sepal.Width~Species,var.equal=F,data=.)

iris %>%
  filter(Species!='versicolor') %>%
  t.test(Sepal.Width~Species,var.equal=F,data=.)

iris %>%
  filter(Species!='setosa') %>%
  t.test(Sepal.Width~Species,var.equal=F,data=.)

```

But what's wrong with this? Since each test has an alpha of 0.05, we have a potential overall alpha of 0.15. That's a 15% chance of Type I error. To control this we'll use an overall test. But don't throw out the pairwise t-tests, they'll be use as post-hoc tests if the overall test is significant.

First we need to run a Levene's test for equality of variances. 

Look at the standard deviations of Sepal.Width for each of the three species.
```{r}
sd(Sepal.Width~Species, data=iris)
```
then examine the levene's test for eaual variances
```{r}
library(RcmdrMisc)
with(iris, leveneTest(Sepal.Width, Species, center=mean))
```

This test has equal variances as the null hypothesis, so we would prefer not to reject the null hypothesis. The p-value of 0.5498 fails to reject $H_0$, so we can go ahead and run the ANOVA including the pairwise comparisons. The pairwise comparisons are simply the t-tests we did above, but run in this context the output is cleaner and we have a couple of extra features.

Note also the use of the aov function. It's almost the same as lm, but has some good features when we're fitting the ANOVA model.
```{r}
AnovaModel.1 <- aov(Sepal.Width ~ Species, data=iris)
summary(AnovaModel.1)


#Alternatively:
AnovaModel.1 <- lm(Sepal.Width ~ Species, data=iris)
anova(AnovaModel.1)
```

The ANOVA table shows a strong rejection of $H_0: \mu_1=\mu_2=\mu_3$, so we look at the pairwise tests (called a Post Hoc analysis) for significant differences. We use the general linear hypotheses function.

```{r}
favstats(Sepal.Width~Species,data=iris)
```

Now do a post-hoc analysis.
```{r}
.Pairs <- glht(AnovaModel.1, linfct = mcp(Species = "Tukey"))
cat('pairwise tests:\n')
summary(.Pairs)
```

```{r}
cat('confidence intervals:\n')
confint(.Pairs)
```

```{r}
cat('compact letter display:\n')
cld(.Pairs)
```

```{r}
old.oma <- par(oma=c(0,5,0,0))
plot(confint(.Pairs))
par(old.oma)
remove(.Pairs)
```


```{r}
with(iris,pairwise.t.test(Petal.Width,Species,p.adjust.method = "bonf"))
```


# The non-parametric one-way test

If we have good reason to suspect non-normality, we can run the equivalent non-parametric test. It's the Kruskal-Wallis test under the Nonparametrics item of Statistics. Note that medians for the groups are automatically compared. If the groups have similar shapes (from the box plots) this is a test for difference in medians.

```{r}
favstats(Sepal.Width~Species, data=iris)
kruskal.test(Sepal.Width ~ Species, data=iris)
```

Now, since we reject $H_0$, run wilcox rank sum tests on pairs of levels to determine which pairs of levels are statistically different. 

Done individually
```{r}
iris %>%
  filter(Species!='virginica') %>%
  wilcox.test(Sepal.Width~Species,var.equal=F,data=.)

iris %>%
  filter(Species!='versicolor') %>%
  wilcox.test(Sepal.Width~Species,var.equal=F,data=.)

iris %>%
  filter(Species!='setosa') %>%
  wilcox.test(Sepal.Width~Species,var.equal=F,data=.)
```

or using pairwise
```{r}
with(iris,
     pairwise.wilcox.test(Sepal.Width,
                          Species,
                          p.adjust.method='none')
     )
```

and using a bonferroni correction
```{r}
with(iris,
     pairwise.wilcox.test(Sepal.Width,
                          Species,
                          p.adjust.method='bonferroni'))
```

And as we saw with the parametric testing, we have strong significance for all three pairs with versicolor and virginica the 'closest' to each other.
