---
title: "Homework #5"
author: "Glenn Clapp"
due: "04 October, 2018"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, message=FALSE,error=FALSE}
library(mosaic)
library(reticulate) # Allows python and R to share namespace
knitr::knit_engines$set(Python = reticulate::eng_python) # Tells the knit function about the python engine
```

7.2.2 (3pts)

```{r}
subjects = data.frame(
  age = c(62,62,68,48,51,60,51,57,57,41,62,50,53,34,62,61)
)

t.test(subjects$age,var.equal=T,mu=60)
favstats(subjects$age)

```
Yes, a 95% confidence interval estimates the true mean to be between 50.21 and 59.67. The entirety of that interval is less than 60.

7.2.4 (3pts)
```{r}
n_visits=25
xbar_visits = 4.8
s_visits = 2

ts <- -qt((1-.95)/2,lower.tail=TRUE,df=n_visits-1)
standard_error = s_visits/n_visits^0.5
print(ts)
#pt(ts,df=n_visits-1)
lower.ci=xbar_visits-ts*standard_error
upper.ci=xbar_visits+ts*standard_error

print(lower.ci)
print(upper.ci)
```

Can it be concluded from these data that the population mean is greater than four visits per patient? NO.

What assumptions are necessary? Random sample selection and normally distributed population.

7.3.2 (3pts)
```{r}
n_fract = 31
n_healthy = 31

xbar_fract = 76.9
s_fract = 12.6

xbar_healthy = 90.9
s_healthy = 12.5

xbar_paired = xbar_healthy - xbar_fract
s_paired = ((s_fract^2+s_healthy^2)/n_fract)^0.5 #Same sample size simplification

standard_error = s_paired/n_fract^0.5
ts <- -qt((1-.95)/2,lower.tail=TRUE,df=n_fract-1)
print(ts)
lower.ci=xbar_paired-ts*standard_error
upper.ci=xbar_paired+ts*standard_error

print(lower.ci)
print(upper.ci)

```

Is the healthy group higher? Yes. The 95% confidence interval in a paired t-test is entirely positive.

7.3.4 (3pts)
```{r}
 hypertension <- read.csv('EXR_C07_S03_04.csv')

t.test(Length~Group,data=hypertension,var.equal=T)
```

Can we conclude that patients with primary hypertension on average havehigher cholesterol levels than normotensive patients?
Yes. The true difference in peans is not equal to 0 p = 0.0001125



7.3.12 (3pts)


Can we conclude that, on the average, lymphocyctes and tumor cells differ in size? Yes.
```{r}
 tumors <- read.csv('EXR_C07_S03_12.csv')

t.test(Size~Group,data=tumors,var.equal=T)
```
7.4.2 (3pts)
```{r}
wellbeing <- read.csv('EXR_C07_S04_02.csv')

t.test(wellbeing$Base-wellbeing$Follow,data=tumors,var.equal=T)
```

Is there sufficient evidence to indicate that quality of physical well-being significantly decreases in the first week of discharge among patients who receive a phone call? No.3

7.9.2 (3pts)

7.10.2 (3pts)

13.4.2 (3pts)
13.6.2 (3pts)
