---
title: "RCBD and 2-way ANOVA examples"
author: "David Zeitler"
date: "`r as.character(Sys.Date())`"
output: pdf_document
---

```{r setup, warning=F, message=F }
library(RcmdrMisc)
library(mosaic)
library(readr)
```

# Example 8.3.1 Randomized Complete Block Design (RCBD)

```{r}
Example8.3.1 <- 
  read_csv("EXA_C08_S03_01.csv")
```

Convert the numeric input to a factor.
```{r}
Example8.3.1 <- mutate(Example8.3.1,
  ageGroup = factor(ageGroup, labels=c('Under 20','20 to 29','30 to 39','40 to 49','50 and over')),
  method = factor(method)
)
```

Run the ANOVA using the linear model command, giving it the model time ~ ageGroup + method.

Note that we didn't check for conditions first. With more advanced
models like this, we use the model residuals to check for conditions
and so need to actually run the model before we can check them.

The exception to the above is running a Levene's test. However we
cannot run a Levene's test with the RCBD design when we have only one
replicate like we have here.

```{r}
LinearModel.1 <- lm(time ~ ageGroup + method, data=Example8.3.1)#+method indicates that method is a blocking factor.
summary(LinearModel.1)
```
```{r}
anova(LinearModel.1)
```

```{r}
LinearModel.1 <- lm(time ~ ageGroup +method, data=Example8.3.1)
anova(LinearModel.1)
```

To check for normality we can look at a normal probability plot of the
residuals from the model. We get this plot from the model menu under graphs.

Recall that the model is
$x_{ij}=\mu+\beta_i+\tau_j+\epsilon_{ij}$ where the error term
$\epsilon_{ij}$ is a zero mean normal random variable. We see here
that while things look pretty good, there are a couple of points that
may indicate some deviation from normality. Since ANOVA is relatively
robust to the normality assumption, we won't worry about this minor
deviation.

```{r}
gf_qq(~residuals(LinearModel.1)) %>%
  gf_qqline()
```
We should also look at the residuals by predictions.
```{r}
gf_point(rstandard(LinearModel.1)~predict(LinearModel.1)) %>%
  gf_hline(yintercept=0)
```

Our conditions appear to be ok for the model, so take a look back at
the Anova table. We see that both ageGroup and method are strongly
significant. The ageGroup factor is our block and having it
significant just tells us that we were right to block our
experiment. Our primary null hypothesis is that the methods are no
different and we see that we should be rejecting it.

Looking at the component residual plots gives us a look at the
response to each factor with the effect of the other removed. We're primarily interessted here in the method plot.

```{r}
crPlots(LinearModel.1, span=0.5)
```

Next we look at effect or means plots by each factor, these are more
common than component + residual plots and are similar to the
confidence interval plots for one-way and the confidence interval
plots shown on page 342 of your text.

```{r}
library(effects)
plot(allEffects(LinearModel.1))
```

Note that we did not do pairwise comparisons. These are possible, but
are more difficult than they were for the one-way model and we have
enough to work on already $\ddot\smile$.

# Example 8.5.2 2-way ANOVA with interaction

```{r}
Example8.5.2 <- 
  read_csv("EXA_C08_S05_02.csv")
```

Convert factor A and B to PatientType and NurseAge with appropriate levels.

Note: I really hate these minitab examples. No names on variables,
hard to read, crap with no excuse... Glad we're working with R.

```{r}
Example8.5.2 <- mutate(Example8.5.2, 
  PatientType = factor(A, labels=c('Cardiac','Cancer','C.V.A.','Tuberculosis')),
  NurseAge    = factor(B, labels=c('20 to 29','30 to 39','40 to 49','50 and over'))
)
```

Look at the boxplots of the response by each factor. These give us an
initial idea of what's going on. Note the potential problem with
equality of variance with respect to nurseAge.

```{r}
gf_boxplot(HOME~NurseAge, data=Example8.5.2)
```

```{r}
gf_boxplot(HOME~PatientType, data=Example8.5.2)
```

The full box plots for the model with interaction would use plots for each of the 16 cells.
```{r}
gf_boxplot(HOME~PatientType:NurseAge,data=Example8.5.2,
       scales=list(x=list(rot=90)))
```

Let's run a Levene's test here. We get a p-value of 0.2534, so we don't reject the hypothesis of equal variances. We'll keep going.

```{r}
leveneTest(HOME ~ NurseAge*PatientType, data=Example8.5.2, center="mean")
```

Run the model using the linear model procedure and get an Anova table again. Note the '*' in the model formula. This gives us both the main effects and interaction term.

```{r}
LinearModel.2 <- lm(HOME ~ NurseAge*PatientType, data=Example8.5.2)
anova(LinearModel.2)
```
```{r}
summary(LinearModel.2)
```
```{r}
gf_boxplot(HOME~PatientType|NurseAge,data=Example8.5.2,
       scales=list(x=list(rot=90)))
gf_boxplot(HOME~NurseAge|PatientType,data=Example8.5.2,
       scales=list(x=list(rot=90)))
```

This Anova table is slightly more complex. Along with the main effects
NurseAge and PatientType, we have a new line labeled
NurseAge:PatientType. This is the interaction term that captures
different responses to PatientType depending on the NurseAge
category. In particular this interaction is significant. We would
*much* rather it not be significant. Many statisticians suggest
stopping at this point and not looking at main effects since they
cannot be considered individually with this interaction
significant. What this really means is that we cannot look at the
length of home visit for different types of patients without first
considering the age of the nurse (or the other way around). This is a
complex model.

Let's finish looking at our conditions.
```{r fig.height=8}
gf_point(residuals(LinearModel.2)~predict(LinearModel.2)) %>% gf_hline(yintercept = 0)
```

```{r}
gf_qq(~residuals(LinearModel.2)) %>% gf_qqline()
```

We see variance 'fan out' in the residual plot and possible normality problems in the high end of the qq-plot. We will accept this for now, but make note of it.

Now let's take a look at some means plots. We didn't do this for the
RCBD analysis, but it's important here. It gives us a good picture of
the interaction that's going on. We're choosing to see the plots with
95% confidence intervals about the means.

```{r}
with(Example8.5.2, plotMeans(HOME, NurseAge, PatientType, error.bars="conf.int", level=0.95))
with(Example8.5.2, plotMeans(HOME, PatientType, NurseAge, error.bars="conf.int", level=0.95))
```

This kind of plot takes a bit of time to absorb. First note that we
have 4 age groups across the x axis. The y axis is the mean of the
length of home visit (response). There are 4 different lines with
different colors and symbols, one for each patient type. The
interaction we're seeing comes from the fact that those four lines are
not even close to parallel and sometimes cross. 
