---
title: "Sampling Distributions for mean"
author: "D. Zeitler"
date: "September 14, 2017"
output: pdf_document
---

```{r setup,message=FALSE,warning=FALSE}
library(mosaic)
library(ggformula)
```

# Take samples manually first.
For each, look at the rectangles page and write down the size of each of the rectangles for n=5 and n=25.
Take the mean and standard deviation of each and add it to the google worksheet. Use R to do your calculations.
```{r}
cat('Sample 5 rectangles:\n')
sample(1:100,5,replace = T)
cat('Sample 25 rectangles:\n')
sample(1:100,25,replace = T)
```
```{r}
s11 <- c(1,18,24,7)
favstats(s11)
```

Once the data file has been saved, download it to read in here.
```{r}
RectangleStats <- read.csv('RectangleStats.csv')
RectangleStats
```

Now let's do some calculations. Theoretically the mean of this distribution is 6.26 and the standard deviation is 5.69.
```{r}
5.69/sqrt(5)
5.69/sqrt(25)
```

```{r}
RectangleStats %>% summary()
```
```{r}
gf_qq(~Mean5,data=RectangleStats)
gf_qq(~Mean25,data=RectangleStats)
```


We have the population, so let's see what the mean ($\mu$) and standard deviation ($\sigma$) are.

```{r}
Rectangles <- read.csv('Rectangles.csv')
histogram(~Area,data=Rectangles,breaks = 25)
```

We can now calculate the mean and standard deviation from the population.

* The mean is $\mu$ = `r mean(~Area,data=Rectangles)`.
* And the standard deviation is $\sigma$ = `r sd(~Area,data=Rectangles)`

```{r}
favstats(~Area,data=Rectangles)
```

# Sampling distribution

So now how does the sampling distribution of the mean look? We'll simulate many samples and see.
```{r}
#sim.1 takes 1000 samples of size 5.If the size were 1, the sample mean distribution would look very much
#like the parent distribution.

sim.1 <- do(1000) * {Rectangles %>% sample_n(5,replace = T) %>% favstats(~Area,data=.)}
histogram(~mean,data=sim.1,breaks = 25)
densityplot(~mean,data=sim.1)
histogram(~sd,data=sim.1)
densityplot(~sd,data=sim.1)
```

```{r}
sim.1 <- do(100) * {Rectangles %>% sample_n(25) %>% favstats(~Area,data=.)}
histogram(~mean,data=sim.1)
densityplot(~mean,data=sim.1)
histogram(~sd,data=sim.1)
densityplot(~sd,data=sim.1)
```
