---
title: "Homework 8"
author: "Glenn Clapp"
date: "`r as.character(Sys.Date())`"
output: pdf_document
---

```{r echo=FALSE, message=F, warning=F}
# include this code chunk as-is to set options
knitr::opts_chunk$set(comment=NA, prompt=TRUE)
options(xtable.comment = FALSE)
library(RcmdrMisc)
library(xtable)
library(MASS)
library(mosaic)
library(readr)
```

This function will be used to analyze regression models throughout this homework.
```{r}
residual.analysis <- function(model) {
  print( xyplot(rstandard(model)~predict(model)) )
  print( qqmath(resid(model),
                panel=function(...){
                  panel.qqmath(...)
                  panel.qqmathline(resid(model))
                })
  )
}
```
9.3.4 (2pts)

Reiss et al. compared point-of-care and standard hospital laboratory assays for monitoring patients receiving a single anticoagulant or a regimen consisting of a combination of anticoagulants. It is quite common when comparing two measuring techniques, to use regression analysis in which one variable is used to predict another. In the present study, the researchers obtained measures of international normalized ratio (INR) by assay of capillary and venous blood samples collected from 90 subjects taking warfarin. INR, used especially when patients are receiving warfarin, measures the clotting ability of the blood. Point-of-care testing for INR was conducted with the CoaguChek assay product. Hospital testing was done with standard hospital laboratory assays. The authors used the hospital assay INR level to predict the CoaguChek INR level. The meaurements are given in the following table.

```{r}
Blood = read_csv("EXR_C09_S03_04.csv")
scatterplot(Y~X, data=Blood, boxplots=F )
Regression.Model <- lm(Y~X, data=Blood)
summary(Regression.Model)
residual.analysis(Regression.Model)

coef(Regression.Model)
confint(Regression.Model) #These are confidence intervals on the coeefficients of the model
```

9.4.2 (6pts)

A. Compute the coefficient of determination. 0.8625108
B. Prepare an ANOVA table and use the F statistic to test the null hypothesis that $\beta_1$ = 0. Let $\alpha$ = 0.05.
The F statistic is 92.41
C. Use the t statistic to test the null hypothesis that $\beta_1$ = 0 at the 0.05 level of significance.
The p value is 2.24e-15 = 0
D. Determine the p value for each hypothesis test.
E. State your conclusions in terms of the problem. There is sufficient evidence to conclude that $\beta_1$ is not equal to 0.
F. Construct the 95 percent confidence interval for $\beta_1$. 0.6842 - 1.0408

9.5.2(2pts)
A. for the value of x indicated, A. contruct the 95 percent confidence interval for $\mu_y|x$ and B. construct the 95 percent prediction interval for Y.

Refer to Exercise 9.3.4 and let X = 1.6
```{r}
newdata = data.frame(X=1.6) 
# This creates a point at which we're interested in the intervals. Note that a 
# list of points could be defined with newdata = data.frame(X=c(1,2,3))
predict(Regression.Model,newdata,interval = 'confidence')
predict(Regression.Model,newdata,interval='predict')
```
Chapter 9 review questions:
16 (5pts):
A study by Scrogin et al. was designed to assess the effects of concurrent manipulations of dietary NaCl and calcium on blood pressure as well as blood pressure and catecholamine responses to stress. Subjects were salt-sensitive, spontaneously hypertensive male rats. Among the analyses performed by the investigators was a correlation between baseline blood pressure and plasma epinephrine concentration (E). The following data on these two variables were collected. Let $\alpha$ = 0.01.
```{r}
REV16 = read_csv("REV_C09_16.csv")
scatterplot(PLASMAE~BP, data=REV16, boxplots=F )
Regression.Model.REV16 <- lm(PLASMAE~BP, data=REV16)
summary(Regression.Model.REV16)
residual.analysis(Regression.Model.REV16)

coef(Regression.Model.REV16)
confint(Regression.Model.REV16,level = 0.99) #These are confidence intervals on the coeefficients of the model
```

There is enough evidence to conclude that there is a positive relationship between blood pressure and plasma. The confidence interval (alpha=0.01) is entirely above zero.

18 (5pts): Run both parametric and non-parametric correlations and comment on the differences seen.
Maria Mathias conducted a study of hyperactive children. She measured the children's attitude, hyperactivity, and social behavior before and after treatment. The following table shows the 31 subjects the age and improvement scores from pre-treatment to post-treatment for attitude (ATT), social behavior (SOC), and hyperactivity (HYP). A negative score for HYP indicates an improvemnent in hyperactivity; a positive score in ATT or SOC indicates improvement. Perform an analysis to determine if there is evidence to indicate that age (years) is correlated with any of the three outcome variables. Let $\alpha$ = 0.05 for all tests.

```{r}
CHILD = read_csv("REV_C09_18.csv")
scatterplot(ATT~AGE, data=CHILD, boxplots=F )
scatterplot(HYP~AGE, data=CHILD, boxplots=F )
scatterplot(SOC~AGE, data=CHILD, boxplots=F )

Regression.Model.ATT <- lm(ATT~AGE, data=CHILD)
Regression.Model.HYP <- lm(ATT~HYP, data=CHILD)
Regression.Model.SOC <- lm(ATT~SOC, data=CHILD)

summary(Regression.Model.ATT)
summary(Regression.Model.HYP)
summary(Regression.Model.SOC)

residual.analysis(Regression.Model.ATT)
residual.analysis(Regression.Model.HYP)
residual.analysis(Regression.Model.SOC)

coef(Regression.Model.ATT)
coef(Regression.Model.HYP)
coef(Regression.Model.SOC)

confint(Regression.Model.ATT,level = 0.95) #These are confidence intervals on the coeefficients of the model
confint(Regression.Model.HYP,level = 0.95)
confint(Regression.Model.SOC,level = 0.95)
```

There is sufficient evidence to conclude that there is a relationship between attitude and age. However, there was not sufficient evidence to conclude that there was a relationship between either of the other characteristics and age at the 0.05 level. The following code block executes the non-parametric tests. For the Spearman test, the critical characteristic below which the null will be rejected is 0.3059 (n=31, alpha=0.05). This value is obtained from table P in appendix A.

```{r}
cat('Attitude w/ pearson:'); cor(ATT~AGE,data=CHILD,method='pearson')
cat('Hyperactivity w/ pearson:'); cor(HYP~AGE,data=CHILD,method='pearson')
cat('Social behavior w/ pearson:'); cor(SOC~AGE,data=CHILD,method='pearson')

cat('\nAttitude w/ spearman:'); cor(ATT~AGE,data=CHILD,method='spearman')
cat('Hyperactivity w/ spearman:'); cor(HYP~AGE,data=CHILD,method='spearman')
cat('Social behavior w/ spearman:'); cor(SOC~AGE,data=CHILD,method='spearman')
```

Using the spearman test, both hyperactivity and social behavior had test statistics below the critical value and are considered significant. Interestingly, this result is flipped from the parametric test for every measure.


20 (5pts)
The following are the pulmonary blood flow (PBF) and pulmonary blood volume (PBV) values recorded for 16 infants and children with congenital heart desease. Find the regression equation describing the linear relationship between the two variables, compute $r^2$, and test $H_0:\beta = 0$ by both the F test and the t test. Let $\alpha$ = 0.05.
```{r}
Pulmonary = read_csv("REV_C09_20.csv")

Regression.Model.PB <- lm(PBF~PBV,data=Pulmonary)
summary(Regression.Model.PB)
residual.analysis(Regression.Model.PB)

```

The equation is PBF = 0.025PBV -0.028; $r^2$ = 0.3937.
The F and p value are 9.091 and 0.0093 respectively at the $\alpha = 0.05$ level of significance which means that there is sufficient evidence to reject the null hypothesis in this case.

13.10.4 (5pts): Show a scatterplot as well as doing the test.
Refer to Exercise 13.10.3. Nozawa et al. also calculated the Japanese Orthopaedic Association score for measuring back pain (JOA). The results for the 20 subjects along with the duration of follow-up are shown in the following table. The higher the number, the lesser the degree of pain. Can we conclude from these data that in general there is a relationship between length of follow-up and JOA score at the time of the operation? Let $\alpha$ = 0.05.

```{r}
backpain = read_csv("EXR_C13_S10_04.csv")
scatterplot(JOA~MONTHS, data=backpain, boxplots=F )
cat('JOA score w/ spearman non-parametric:'); cor(JOA~MONTHS,data=backpain,method='spearman')
```

There are 20 observations in this data. With an $\alpha = 0.05$ confidence level, and using table P in the appendix, the critical statistic value was found to be 0.3789. The value obtained was lower than this, leading us to reject the null hypothesis that there is no relationship between the JOA score and the duration of follow up.

3 predictors: 10.3.2, 10.4.2, 10.5.2 all use 10.3.2 data (10pts)
Family caregiving of older adults is more common in Korea than in the United States. Son et al. studied 100 caregivers of older adults with dementia in Seoul, South Korea. The dependent variable was caregiver burden as measured by the Korean Burden Inventory (KBI). Scores ranged from 28 to 140, with higher scores indicating higher burden. Explanatory variables were indexes that measured the following:

ADL: Total activites of daily living (low scores indicate that the elderly perform activities independently).

MEM: Memory and behavioral problems (higher scores indicate more problems).

COG: Cognitive impairment (lower scores indicate a greater degree of cognitive impairment).


```{r}
care <- 
  read.csv("EXR_C10_S03_02.csv")

scatterplotMatrix(~KBI+ADL+MEM+COG, smooth=FALSE, 
  diagonal = 'density', data=care)
RegModel.care <- lm(KBI~ADL+MEM+COG, data=care)
residual.analysis(RegModel.care)
summary(RegModel.care)

cor(care[,c("KBI","ADL","MEM","COG")], use="complete")
partial.cor(care[,c("KBI","ADL","MEM","COG")], use="complete")
```

```{r}
newdata = data.frame(ADL=c(95), MEM=c(35), COG=c(0))
# This creates a point at which we're interested in the intervals. Note that a 
# list of points could be defined with newdata = data.frame(X=c(1,2,3))
predict(RegModel.care,newdata,interval = 'confidence')
predict(RegModel.care,newdata,interval='predict')
```
R insisted that I have a value for all three independent variables to construct intervals. Only two are given in the problem statement. I expect that one of the three 'COG' is not significant and is to be excluded. I'm not sure how to do this in R.

A. Calculate the coefficient of multiple determination
B. Perform an analysis of variance
C. Test the significance of each $\beta_i$(i>0). Let $\alpha$ = 0.05 for all tests of significance and determine the p value for all tests

D. Construct a 95 percent confidence interval for each significant sample slope.

E. Compute the y value and construct a 95 percent confidence and 95 percent prediction interval for $x_1j$=95 and $x_2j$=35. 

6 predictors: 10.3.6, 10.4.6, 10.5.6 All use 10.3.6 data (10pts)
The following data were collected on a simple random sample of 20 patients with hypertension. The variables are
Y = mean arterial blood pressure (mmHg)
$X_1$=age(years)
$X_2$=weight(kg)
$X_3$=body surface area($m^2$)
$X_4$=duration of hypertension(years)
$X_5$=basal bulse(bpm)
$X_6$=measure of stress
A. Calcultate the coefficient of multiple determination
B. Perform an analysis of variance
C. Test the significance of each $\beta_i$(i>0). Let $\alpha$ = 0.05 for all tests of significance and determine the p value for all tests
D. Construct a 95 percent confidence interval for each significant sample slope.
E. Compute the y value and construct a 95 percent confidence and 95 percent prediction interval for 
$x_1j$=50
$x_2j$=95
$x_3j$=2.00
$x_4j$=6.00
$x_5j$=75
$x_6j$=70
```{r}
hypertension <- 
  read.csv("EXR_C10_S03_06.csv")

scatterplotMatrix(~Y+X1+X2+X3+X4+X5+X6, smooth=FALSE, 
  diagonal = 'density', data=hypertension)
RegModel.hypertension <- lm(Y~X1+X2+X3+X4+X5+X6, data=hypertension)
residual.analysis(RegModel.hypertension)
summary(RegModel.hypertension)

cor(hypertension[,c("X1","X2","X3","X4","X5","X6")], use="complete")
partial.cor(hypertension[,c("X1","X2","X3","X4","X5","X6")], use="complete")

newdata = data.frame(X1=c(50),X2=c(95),X3=c(2.00),X4=c(6.00),X5=c(75),X6=c(70))
# This creates a point at which we're interested in the intervals. Note that a 
# list of points could be defined with newdata = data.frame(X=c(1,2,3))
predict(RegModel.hypertension,newdata,interval = 'confidence')
predict(RegModel.hypertension,newdata,interval='predict')
```

I think that his problem also suffers from the fact that I'm unsure how to eliminate factors that are not significant. I think that would substantially change my results.
