---
title: "Homework 2"
author: "Glenn Clapp"
due: "13 September, 2018"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, message=FALSE,error=FALSE}
library(tidyverse)
library(ggformula)
library(mosaic)
library(readr)
library(moments)
```

# Exercise 4.4.4: What is the probability that an individual selected at random used:
# A. Five addictive substances?
0.0501

# B. Fewer than three addictive substances?
0.6255 = 0.1853+0.4402

# C. More than six addictive substances?
0.0232 = 0.0077+0.0116+0.0026+0.0013

# D. Between two and five adivctive substances, inclusive?
0.7659 = 0.4402+0.1828+0.0927+0.0502
```{r}
substance_use_data = data.frame(
  frequency = c(144,342,142,72,39,20,6,9,2,1)
  )
probability = substance_use_data$frequency/777
substance_use_data <- cbind(substance_use_data,probability)
#plot(substance_use_data$probability)
substance_use_data
```
# Exercise 4.6.2: Given the standard normal distribution find the probability that a z picked at random will have a value between z = -2.87 and z = 2.64.

```{r}
pnorm(2.64) - pnorm(-2.87)
```
# Exercise 4.6.4: P(z>= -0.55)
```{r}
1.0 - pnorm(-0.55)
```
# Exercise 4.6.6: P(z<2.33)
```{r}
pnorm(2.33)
```
# Exercise 4.6.8: P(-2.58<=z<=2.58)
```{r}
pnorm(2.58) - pnorm(-2.58)
```
# Exercise 4.6.10: P(z=0.74)
```{r}
dnorm(0.74,0,1)
```
# Exercise 4.6.12: P(-2.67<=z<=z) = 0.9718
```{r}
qnorm(0.9718+pnorm(-2.67))
```

# Exercise 4.6.14: P(z<=z<=2.98) = 0.1117
```{r}
qnorm(pnorm(2.98)-0.1117)
```
# Exercise 4.7.2
```{r}
MEAN = 140
SD = 50
A = pnorm(200, lower.tail = F, mean=MEAN, sd=SD)
B = pnorm(100, lower.tail = T, mean=MEAN, sd=SD)

print(A)
print(B)
```
# Exercise 4.7.6
```{r}
MEAN = 75
SD = 25 #variance = 625

#A = P(50<=x<=100)
A = pnorm(100, lower.tail = T, mean=MEAN, sd=SD)-pnorm(50, lower.tail = T, mean=MEAN, sd=SD)

#B = P(x>90)
B = pnorm(90, lower.tail = F, mean=MEAN, sd=SD)

#C = P(x<60)
C = pnorm(60, lower.tail = T, mean=MEAN, sd=SD)

#D = P(x>=85)
D = pnorm(85, lower.tail = F, mean=MEAN, sd=SD)

#E = P(30<=x<=110)
E = pnorm(110, lower.tail = T, mean=MEAN, sd=SD)-pnorm(30, lower.tail = T, mean=MEAN, sd=SD)

print(A)
print(B)
print(C)
print(D)
print(E)
```
# Review question 18 (Assume sd = 1.0)

```{r}
m = 2.0
SD = 1
#A = dnorm(3,m,sd) Why won't dnorm knit with non-numeric arguments? error on knitting.
A = dnorm(3,2.0,1)
B = pnorm(2,lower.tail = T, mean = m, sd = SD)
C = pnorm(5,lower.tail = T, mean = m, sd = SD) - pnorm(3,lower.tail = T, mean = m, sd = SD)

print(A)
print(B)
print(C)
```
# Review question 22 
```{r}
m = 10
SD = 3
A = pnorm(4,lower.tail = T, mean = m, sd = SD)
B = pnorm(5,lower.tail = F, mean = m, sd = SD)
C = pnorm(3,lower.tail = T, mean = m, sd = SD)
  
print(A)
print(B)
print(C)
```
# Review question 26
```{r}
P = 1-0.754
P_onetail = P/2
z = -1 * qnorm(P_onetail)
print(paste("k = ", z))
```
# Review question 28
```{r}
P = 0.008
SD = 10
X=40 #P(x<=40)
z = qnorm(P)
# z = (x-u)/SD
u = (z*SD - X)*-1
print(u)
```

# Review question 30
```{r}
P = 0.0526
SD = 5
X=25 #P(x>=25)
z = qnorm(P,lower.tail = F)
# z = (x-u)/SD
u = (z*SD - X)*-1
print(u)
```

# Review question 32
```{r}
P = 0.9772
u = 30
X = 50 #P(x<=50)
z = qnorm(P, lower.tail = T)
#z = (x-u)/SD
SD = (X-u)/z
print(SD)
```

